Interpolation of Sparse High-Dimensional Data\thanks{This projected was supported by NSF grant}

### 2019-02-08 Documents_[2019]/approximation/paper.tex
Ready to run approximation experiments.

\documentclass[sigconf]{acmart}

\author{Thomas C. H. Lux}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \department{Dept. of Computer Science}
  \institution{Virginia Polytechnic Institute and State University}
  \city{Blacksburg} 
  \state{Virginia} 
  \postcode{24060}
}
\email{tchlux@vt.edu}

\author{Layne T. Watson}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \department{Depts. of Computer Science, Mathematics, and Aerospace \& Ocean Engineering}
  \institution{Virginia Polytechnic Institute and State University}
}

\author{Tyler H. Chang}
\affiliation{%
  \department{Dept. of Computer Science}
  \institution{Virginia Polytechnic Institute and State University}
}

% The default list of authors is too long for headers.
\renewcommand{\shortauthors}{T. Lux et al.}
\def\mn#1{\left\Vert #1\right\Vert_\infty}

\def\Bigl{\mathopen\Big}
\def\Bigr{\mathclose\Big}

\title{Interpolation of Sparse High-Dimensional Data*}

%% \documentclass[letterpaper, 10 pt, conference]{ieeeconf}  % Comment this line out
%%                                                           % if you need a4paper
%% %\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
%% \IEEEoverridecommandlockouts                              % This command is only needed 
%% %% % copyright info
%% %% \IEEEpubid{\makebox[\columnwidth]{978-1-5386-6133-8/18/\$31.00
%% %%         ~\copyright~2018 IEEE \hfill}\hspace{\columnsep}
%% %%         \makebox[\columnwidth]{ }}

%% % if you want to use the \thanks command
%% \overrideIEEEmargins
%% % See the \addtolength command later in the file to balance the column lengths on the last page of the document

%% \title{\LARGE \bf Interpolants Scalable by Dimension* }

%% \author{Thomas C. H. Lux$^{1}$, Layne T. Watson$^{2}$, Tyler H. Chang$^{1}$% <-this % stops a space
%% \thanks{*This work was supported by the National Science Foundation Grant CNS-1565314.}%
%% \thanks{$^{1}$Doctoral student, Department of Computer Science, Virginia Polytechnic Institute and State University, Blacksburg, Virginia 24060, USA {\tt\small (tchlux at vt.edu)}}%
%% \thanks{$^{2}$Department of Computer Science, Virginia Polytechnic Institute and State University, Blacksburg, Virginia 24060, USA}%
%% }

% The following packages can be found on http:\\www.ctan.org
\usepackage{tikz}     % For adding axes to plots
\usetikzlibrary{positioning}
\usepackage{graphicx} % For including PDFs
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{booktabs} % For formal tables
\usepackage{hyphenat} % custom hyphenation rules
\hyphenation{op-tical net-works semi-conduc-tor}

%% \usepackage{color}   % Needed to color text for drafts
%% \usepackage{algorithm}     % needed for the algorithm block
%% \usepackage{algpseudocode} % algorithmicx psuedo-code styling
%% \usepackage[pdftex,colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black]{hyperref}
%% \usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black]{hyperref}

@@

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}
%%====================================================================
%%                    BODY OF PAPER
%%====================================================================

\section*{Abstract}

A rapid increase in the quantity of data available is allowing all fields of science to generate more accurate models of multivariate phenomena. Regression and interpolation become challenging when the dimension of data is large, especially while maintaining tractable computational complexity. While regression is a popular approach to solving approximation problems with high dimension, there are often advantages to interpolation. This paper presents error bounds for interpolants on moderately high dimensional problems and contrasts their performance with that of popular regression techniques. Empirical results demonstrate the viability of interpolants for high dimensional approximation problems, suggesting that these techniques are capable of effectively modeling multivariate phenomena while maintaining flexibility in different application domains.

\textbf{Keywords:} Approximation, regression, interpolation, high dimension, error bound

@@

\section{Multivariate Regression}
Multivariate regressors are capable of accurately modeling a complex dependence of a response (in $Y$) on multiple variables (represented as a points in $\mathbb{R}^{d}$). The approximations to some (unknown) underlying function $f: \mathbb{R}^d \rightarrow Y$ are chosen to minimize some error measure related to data samples $f\bigl(x^{(i)}\bigr)$. For example, least squares regression uses the sum of squared differences between modeled function values and true function values as an error measure. In this section and the next, some techniques are limited to approximating real valued functions ($Y \subset \Re$). These techniques are extended to real vector-valued functions by repeating the construction for each component of the vector outputs. Throughout the following, $x$ denotes a $d$-tuple, $x_i$ the $i$th component of $x$, and $x^{(i)}$ the $i$th $d$-tuple data point. Different symbols are used to represent the approximation function $\hat f$.

@@

A well-studied technique for classification and approximation is the nearest neighbor algorithm \cite{cover1967nearest}. Nearest neighbor inherently utilizes the convex region $C^{x^{(i)}}$ (Voronoi cell \cite{dirichlet1850reduction}) consisting of all points closer to $x^{(i)}$ than to any other point $x^{(j)}$. The Voronoi mesh smooths the nearest neighbor approximation by utilizing the Voronoi cells to define support via a generic basis function $v: \mathbb{R}^d \rightarrow \mathbb{R}_+$ given by
$$ v^{x^{(i)}}(y) = \left(1 - \frac{||y - x^{(i)}||_2}{2 \ h(y - x^{(i)} \mid x^{(i)})} \right)_+, $$
where $h(w \mid x^{(i)})$ is the distance between $x^{(i)}$ and the boundary of the Voronoi cell $C^{x^{(i)}}$ in the direction $w$. $v^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $v^{x^{(i)}}$ has local support, giving the interpolated value
$$ f(y) = \frac{\sum_{i=1}^n v^{x^{(i)}}(y) f(x^{(i)})}{\sum_{i=1}^n v^{x^{(i)}}(y)}, $$
where $0 \leq v^{x^{(i)}}(y) \leq 1$. The computational complexity of interpolation via the Voronoi mesh is $\mathcal{O}(n^2 d)$. All of the approximations are an interpolant involving a convex combination of known function values $f(x^{(i)})$.

\begin{figure}[htb]
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.4\textwidth,]{plot-delaunay-example-prediction.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Throughput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {CDF Value};
  \end{tikzpicture}
  \vspace{-0.3cm}
  \caption{In this HPC I/O example, the general methodology for predicting a CDF and evaluating error can be seen. The Delaunay method chose three source distributions (dotted lines) and assigned weights \{.3, .4, .3\} (top to bottom at arrow). The weighted sum of the three known CDFs produces the predicted CDF (dashed line). The KS Statistic (arrow) computed between the true CDF (solid line) and predicted CDF (dashed line) is 0.2 for this example. The KS test null hypothesis is rejected at $p$-value 0.01, however it is not rejected at $p$-value 0.001.
  \vspace{-.1cm}}
  \label{fig:prediction_example}
\end{figure}

@@

The performance of approximation techniques that predict probability functions can be analyzed through a variety of summary statistics. Mean absolute error, mean squared error, and the max absolute error are all popular measures. This work uses the max absolute error, also known as the Kolmogorov-Smirnov (KS) statistic \cite{lilliefors1967kolmogorov} for its compatibility with the KS test.

The two-sample KS test is a useful nonparametric test for comparing two CDFs while only assuming stationarity, finite mean, and finite variance. The null hypothesis (that two CDFs come from the same underlying distribution) is rejected at level $p \in [0,1]$ when

$$ KS > \sqrt{-\frac{1}{2}\ln\biggl(\frac{p}{2}\biggr)} \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}, $$

with distribution sample sizes $n_1,n_2 \in \mathcal{N}$. For all applications of the KS test presented in this work $n_1 = n_2$. An example of the round-trip prediction methodology from known and predicted distributions to the calculation of error can be seen in Figure \ref{fig:prediction_example}.

@@

\noindent \textbf{\textit{Lemma 1.}} Let $S \subset \mathbb{R}^d$ be open and convex, $\ f: \mathbb{R}^d \rightarrow \mathbb{R},$ and $\nabla f \in Lip_{(\gamma,\|\cdot\|_2)}(S)$, the set of $\gamma$-Lipschitz continuous functions in the $2$-norm. Then for all $x,y \in S$

$$\big|\ f(y) - f(x) - \langle \nabla f(x), y - x \rangle \big| \leq \frac{\gamma \|y - x\|_2^2}{2}.$$

\noindent \textit{Proof.} Consider the function $g(t) = f \big((1-t) x + t y \big),$ $0 \leq t \leq 1,$ whose derivative $g'(t) = \big\langle \nabla f \big((1-t) x + t y \big), y - x \big\rangle$ is the directional derivative of $f$ in the direction $(y - x).$

\begin{align*}
  \big|\ f(y) - f&(x) - \langle \nabla f(x), y - x \rangle \big| &\\ 
     &= \big|\ g(1) - g(0) - g'(0) \big| & \\
     &= \bigg| \int_0^1 g'(t) - g'(0)\ dt \bigg| \\
     &\leq \int_0^1 \big|g'(t) - g'(0)\big|\ dt \\
     &= \int_0^1 \bigg| \big \langle \nabla f\big((1-t)x + ty\big) - \nabla f(x), y - x \big \rangle \bigg|\ dt \\
     &\leq \int_0^1 \big \| \nabla f\big((1-t)x + ty\big) - \nabla f(x) \big \|_2\ \| y - x \|_2\ dt \\
     &\leq \int_0^1 \big ( \gamma\ \|y-x\|_2 \big) \ \big( \|y-x\|_2 \big) t\ dt \\
     &= \frac{\gamma \|y - x\|_2^2}{2}.
\end{align*}

$\qed$

\noindent \textbf{\textit{Lemma 2.}} Let $x, y, v_i \in \mathbb{R}^d,$ $c_i \in \mathbb{R},$ and $|\langle y - x, v_i \rangle| \leq c_i$ for $i = 1,$ $\ldots,$ $d.$ If $M = (v_1,$ $\ldots,$ $v_d)$ is nonsingular, then

$$\|y - x\|_2^2 \leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2,$$

where $\sigma_d$ is the smallest singular value of $M.$

\noindent \textit{Proof.} Using the facts that $M$ and $M^t$ have the same singular values, and $\|M^tw\|_2 \geq \sigma_d \|w\|_2,$ gives

\begin{align*}
     \|y - x\|_2^2 &\leq \frac{\|M^t (y - x)\|_2^2}{\sigma_d^2} \\
                   &=    \frac{1}{\sigma_d^2} \sum_{i=1}^d \langle y - x, v_i \rangle^2 \\
                   &\leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2.
\end{align*}

$\qed$

\noindent \textbf{\textit{Lemma 3.}} Given $\ f,$ $\gamma,$ $S$ as in {\it Lemma 1}, let $X = \{x_0,$ $x_1,$ $\ldots,$ $x_d\}$ $\subset S$ be the vertices of a $d$-simplex, and let $\hat f(x) = \langle c, x - x_0 \rangle + f(x_0),$ $c \in \mathbb{R}^d$ be the linear function interpolating $\ f$ on $X.$ Let $\sigma_d$ be the smallest singular value of the matrix $M = (x_1 - x_0,$ $\ldots,$ $x_d - x_0),$ and $k = \max\limits_{1\ \leq\ j\ \leq\ d} \|x_j - x_0\|_2.$ Then

$$\big\|\nabla f(x_0) - c\big\|_2 \leq \sqrt{d} \frac{\gamma k^2}{\sigma_d}.$$

\noindent \textit{Proof.} Consider $f(x) - \hat f(x)$ along the line segment $z(t) = (1-t)x_0 + t x_j,$ $0 \leq t \leq 1.$ By Rolle's Theorem, for some $0 < \hat t < 1,$ $\big\langle \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0 \big\rangle = 0.$ Now

\begin{align*}
       \big| \big< \nabla f(x_0) &- c, x_j - x_0 \big \rangle \big| \\
        &= \big| \big< \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) + \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0 \big \rangle \big| \\
	&= \big| \big\langle \nabla f(x_0) - \nabla f \big(z(\ \hat t\ )\big), x_j - x_0 \big \rangle \big| \\
	&\leq \big \| \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) \big \|_2 \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_0 - z(\ \hat t\ )\|_2\ \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_j - x_0\|_2^2 \leq \gamma k^2, \\
\end{align*}

for all $1 \leq j \leq d.$ Using {\it Lemma 2},

$$\big \| \nabla f(x_i) - c \big \|_2^2 \leq \frac{d}{\sigma_d^2} \big( \gamma k^2\big)^2 \implies \big \| \nabla f(x_i) - c \big \|_2 \leq \sqrt{d} \frac{\gamma k^2}{\sigma_d}.$$

$\qed$

\noindent \textbf{\textit{Theorem.}} Under the assumptions of {\it Lemma 1} and {\it Lemma 3}, for $z \in S,$

$$ \big|\ f(z) - \hat f(z)\big| \leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d} \frac{\gamma k^2}{\sigma_d} \|z - x_0\|_2.$$

\noindent \textit{Proof.} Let $v = \nabla f(x_0) - c,$ where $\|v\|_2 \leq \sqrt{d} \gamma k^2 / \sigma_d$ by {\it Lemma 3}. Now

\begin{align*}
       \big|\ f(z) - \hat f(z)\ \big| &= \big|\ f(z) - f(x_0) - \langle c, z - x_0 \rangle \big| \\ 
       &= \big|\ f(z) - f(x_0) - \langle \nabla f(x_0) - v, z - x_0 \rangle \big| \\
       &= \big|\ f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle + \langle v , z - x_0 \rangle \big| \\
       &\leq \big|\ f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \big| \langle v , z - x_0 \rangle \big| \\
       &\leq \big|\ f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \|v\|_2 \ \|z - x_0\|_2 \\
       &\leq \big|\ f(z) - f(x_0) - \langle \nabla f(x_0), z - x_0 \rangle \big| + \textstyle{\frac{\gamma k^2 \sqrt{d}}{\sigma_d}} \|z - x_0\|_2 \\
       &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d}\frac{\gamma k^2}{\sigma_d} \|z - x_0\|_2,
\end{align*}

where the last inequality follows from {\it Lemma 1}.

$\qed$

@@

\begin{figure}
  \includegraphics[width=.5\textwidth,height=2cm]{p-hist.pdf}
  \includegraphics[width=.5\textwidth,height=2cm]{f-hist.pdf}
  \includegraphics[width=.5\textwidth,height=2cm]{h-hist.pdf}
  \caption{Histograms of Parkinsons (total UPDRS), forest fire (area), and HPC I/O (mean throughput) response values respectively. Notice that both the forest fire and HPC I/O data sets are heavily skewed.
  \vspace{-.5cm}}
  \label{fig_response_hists}
\end{figure}

@@

\begin{figure*}[htb]
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.95\textwidth,height=4cm]{eval-times.pdf}};
    \node[below=of img, node distance=0cm, yshift=1cm] {Relative Error Tolerance while Bootstrapping};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Fit Time};
  \end{tikzpicture}
  \caption{Time required to generate model fits for each technique with varying relative error tolerance during bootstrapping.
    \vspace{-.3cm}}
  \label{fig_eval_times}
\end{figure*}

@@

\begin{figure*}
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.95\textwidth,height=9.5cm]{all-performance.pdf}};
    \node[below=of img, node distance=0cm, yshift=1cm] {Relative Error Tolerance while Bootstrapping};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Signed Relative Error};
  \end{tikzpicture}
  \caption{The performance of all three techniques with varied relative error tolerance for the bootstrapping parameter. The columns are for Max Box Mesh, Iterative Box Mesh, and Voronoi Mesh, respectively. The rows are for HPC I/O, Forest Fire, and Parkinson's respectively. Notice the techniques' behavior on the Parkinson's and Forest Fire data sets, performance increases with larger error tolerance.}
  \label{fig_all_performance}
\end{figure*}

>>>>

\RequirePackage{fix-cm}
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\smartqed  % flush right qed marks, e.g. at end of proof
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
\usepackage{tikz}     % For adding axes to plots
\usetikzlibrary{positioning}
\usepackage{graphicx} % For including PDFs
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{booktabs} % For formal tables

%
% please place your own definitions here and don't use \def but
% \newcommand{}{}

@@

%% \newcommand{\Bigl}{\mathopen\Big}
%% \newcommand{\Bigr}{\mathclose\Big}
%
% Insert the name of "your journal" with
\journalname{Numerical Algorithms}
%
\begin{document}

\title{Interpolation of Sparse High-Dimensional Data\thanks{This
    projected was supported by NSF grant}
}
% \subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Thomas C. H. Lux   \and
  Layne T. Watson          \and
  Tyler H. Chang           \and
  Yili Hong                \and
  Kirk Cameron
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Thomas Lux \at
  Virginia Polytechnic Institute \& State University (VPI\&SU) \\
  Blacksburg, VA 24061 \\
  \email{tchlux@vt.edu}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
A rapid increase in the quantity of data available is allowing all fields of science to generate more accurate models of multivariate phenomena. Regression and interpolation become challenging when the dimension of data is large, especially while maintaining tractable computational complexity. While regression is a popular approach to solving approximation problems with high dimension, there are often advantages to interpolation. This paper presents error bounds for interpolants on moderately high dimensional problems and contrasts their performance with that of popular regression techniques. Empirical results demonstrate the viability of interpolants for high dimensional approximation problems, suggesting that these techniques are capable of effectively modeling multivariate phenomena while maintaining flexibility in different application domains.

\keywords{approximation \and regression \and interpolation \and high dimension \and error bound}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

@@

\section{Multivariate Regression}
Multivariate regressors are capable of accurately modeling a complex dependence of a response (in $Y$) on multiple variables (represented as a points in $\mathbb{R}^{d}$). The approximations to some (unknown) underlying function $f: \mathbb{R}^d \rightarrow Y$ are chosen to minimize some error measure related to data samples $f\bigl(x^{(i)}\bigr)$. For example, least squares regression uses the sum of squared differences between modeled function values and true function values as an error measure. In this section and the next, some techniques are limited to approximating real valued functions ($Y \subset \mathbb{R}$). These techniques are extended to real vector-valued functions by repeating the construction for each component of the vector outputs. Throughout the following, $x$ denotes a $d$-tuple, $x_i$ the $i$th component of $x$, and $x^{(i)}$ the $i$th $d$-tuple data point. Different symbols are used to represent the approximation function $\hat f$.

@@

A well-studied technique for classification and approximation is the nearest neighbor algorithm \cite{cover1967nearest}. Nearest neighbor inherently utilizes the convex region $C^{x^{(i)}}$ (Voronoi cell \cite{dirichlet1850reduction}) consisting of all points closer to $x^{(i)}$ than to any other point $x^{(j)}$. The Voronoi mesh smooths the nearest neighbor approximation by utilizing the Voronoi cells to define support via a generic basis function $v: \mathbb{R}^d \rightarrow \mathbb{R}_+$ given by
$$ v^{x^{(i)}}(y) = \left(1 - \frac{||y - x^{(i)}||_2}{2 \ h(y - x^{(i)} \mid x^{(i)})} \right)_+, $$
where $h(w \mid x^{(i)})$ is the distance between $x^{(i)}$ and the boundary of the Voronoi cell $C^{x^{(i)}}$ in the direction $w$. $v^{x^{(i)}}\bigl(x^{(j)}\bigr) = \delta_{ij}$ and $v^{x^{(i)}}$ has local support, giving the interpolated value
$$ f(y) = \frac{\sum\limits_{i=1}^n v^{x^{(i)}}(y) f(x^{(i)})}{\sum\limits_{i=1}^n v^{x^{(i)}}(y)}, $$
where $0 \leq v^{x^{(i)}}(y) \leq 1$. The computational complexity of interpolation via the Voronoi mesh is $\mathcal{O}(n^2 d)$. All of the approximations are an interpolant involving a convex combination of known function values $f(x^{(i)})$.

\begin{figure}[htb]
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.4\textwidth,]{Figures/plot-delaunay-example-prediction.pdf}};
    \node[below=of img, node distance=1cm, yshift=1cm] {I/O Throughput};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {CDF Value};
  \end{tikzpicture}
  \vspace{-0.3cm}
  \caption{In this HPC I/O example, the general methodology for predicting a CDF and evaluating error can be seen. The Delaunay method chose three source distributions (dotted lines) and assigned weights \{.3, .4, .3\} (top to bottom at arrow). The weighted sum of the three known CDFs produces the predicted CDF (dashed line). The KS Statistic (arrow) computed between the true CDF (solid line) and predicted CDF (dashed line) is 0.2 for this example. The KS test null hypothesis is rejected at $p$-value 0.01, however it is not rejected at $p$-value 0.001.
  \vspace{-.1cm}}
  \label{fig:prediction_example}
\end{figure}

@@

The performance of approximation techniques that predict probability functions can be analyzed through a variety of summary statistics. This work uses the max absolute error, also known as the Kolmogorov-Smirnov (KS) statistic \cite{lilliefors1967kolmogorov} for its compatibility with the KS test.

The two-sample KS test is a useful nonparametric test for comparing two CDFs while only assuming stationarity, finite mean, and finite variance. The null hypothesis (that two CDFs come from the same underlying distribution) is rejected at level $p \in [0,1]$ when
$$ KS > \sqrt{-\frac{1}{2}\ln\biggl(\frac{p}{2}\biggr)} \sqrt{\frac{1}{n_1} + \frac{1}{n_2}}, $$
with distribution sample sizes $n_1,n_2 \in \mathcal{N}$. For all applications of the KS test presented in this work $n_1 = n_2$. An example of the round-trip prediction methodology from known and predicted distributions to the calculation of error can be seen in Figure \ref{fig:prediction_example}.

@@

\noindent \textbf{\textit{Lemma 1.}} Let $S \subset \mathbb{R}^d$ be open and convex, $f: \mathbb{R}^d \rightarrow \mathbb{R},$ and $\nabla f \in Lip_{(\gamma,\|\cdot\|_2)}(S)$, the set of $\gamma$-Lipschitz continuous functions in the $2$-norm. Then for all $x,y \in S$
%
$$\big|f(y) - f(x) - \langle \nabla f(x), y - x \rangle \big| \leq \frac{\gamma \|y - x\|_2^2}{2}.$$

\noindent \textit{Proof.} Consider the function $g(t) = f \big((1-t) x + t y \big),$ $0 \leq t \leq 1,$ whose derivative $g'(t) = \big\langle \nabla f \big((1-t) x + t y \big), y - x \big\rangle$ is the directional derivative of $f$ in the direction $(y - x).$
%
\begin{align*}
  \big|f(y) - f(x) - \langle \nabla f(x), y - x \rangle \big|
     &= \big|g(1) - g(0) - g'(0) \big| & \\
     &= \bigg| \int_0^1 g'(t) - g'(0)\ dt \bigg| \\
     &\leq \int_0^1 \big|g'(t) - g'(0)\big|\ dt \\
     &= \int_0^1 \bigg| \big \langle \nabla f\big((1-t)x + ty\big) - \nabla f(x), y - x \big \rangle \bigg|\ dt \\
     &\leq \int_0^1 \big \| \nabla f\big((1-t)x + ty\big) - \nabla f(x) \big \|_2\ \| y - x \|_2\ dt \\
     &\leq \int_0^1 \big ( \gamma\ \|y-x\|_2 \big) \ \big( \|y-x\|_2 \big) t\ dt \\
     &= \frac{\gamma \|y - x\|_2^2}{2}.
\end{align*}
%
\qed

\noindent \textbf{\textit{Lemma 2.}} Let $x, y, v_i \in \mathbb{R}^d,$ $c_i \in \mathbb{R},$ and $|\langle y - x, v_i \rangle| \leq c_i$ for $i = 1,$ $\ldots,$ $d.$ If $M = (v_1,$ $\ldots,$ $v_d)$ is nonsingular, then
%
$$\|y - x\|_2^2 \leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2,$$
%
where $\sigma_d$ is the smallest singular value of $M.$

\noindent \textit{Proof.} Using the facts that $M$ and $M^t$ have the same singular values, and $\|M^tw\|_2 \geq \sigma_d \|w\|_2,$ gives
%
\begin{align*}
     \|y - x\|_2^2 &\leq \frac{\|M^t (y - x)\|_2^2}{\sigma_d^2} \\
                   &=    \frac{1}{\sigma_d^2} \sum_{i=1}^d \langle y - x, v_i \rangle^2 \\
                   &\leq \frac{1}{\sigma_d^2} \sum_{i=1}^d c_i^2.
\end{align*}
%
\qed

\noindent \textbf{\textit{Lemma 3.}} Given $f,$ $\gamma,$ $S$ as in {\it Lemma 1}, let $X = \{x_0,$ $x_1,$ $\ldots,$ $x_d\}$ $\subset S$ be the vertices of a $d$-simplex, and let $\hat f(x) = \langle c, x - x_0 \rangle + f(x_0),$ $c \in \mathbb{R}^d$ be the linear function interpolating $f$ on $X.$ Let $\sigma_d$ be the smallest singular value of the matrix $M = (x_1 - x_0,$ $\ldots,$ $x_d - x_0),$ and $k = \max\limits_{1\ \leq\ j\ \leq\ d} \|x_j - x_0\|_2.$ Then
%
$$\big\|\nabla f(x_0) - c\big\|_2 \leq \sqrt{d} \frac{\gamma k^2}{\sigma_d}.$$

\noindent \textit{Proof.} Consider $f(x) - \hat f(x)$ along the line segment $z(t) = (1-t)x_0 + t x_j,$ $0 \leq t \leq 1.$ By Rolle's Theorem, for some $0 < \hat t < 1,$ $\big\langle \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0 \big\rangle = 0.$ Now
%
\begin{align*}
  \big| \big< \nabla f(x_0) - c, x_j - x_0 \big \rangle \big|
        &= \big| \big< \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) + \nabla f\big(z(\ \hat t\ )\big) - c, x_j - x_0 \big \rangle \big| \\
	&= \big| \big\langle \nabla f(x_0) - \nabla f \big(z(\ \hat t\ )\big), x_j - x_0 \big \rangle \big| \\
	&\leq \big \| \nabla f(x_0) - \nabla f\big(z(\ \hat t\ )\big) \big \|_2 \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_0 - z(\ \hat t\ )\|_2\ \|x_j - x_0\|_2 \\
	&\leq \gamma \|x_j - x_0\|_2^2 \leq \gamma k^2, \\
\end{align*}
%
for all $1 \leq j \leq d.$ Using {\it Lemma 2},
%
$$\big \| \nabla f(x_i) - c \big \|_2^2 \leq \frac{d}{\sigma_d^2} \big( \gamma k^2\big)^2 \implies \big \| \nabla f(x_i) - c \big \|_2 \leq \sqrt{d} \frac{\gamma k^2}{\sigma_d}.$$
%
\qed

\noindent \textbf{\textit{Theorem.}} Under the assumptions of {\it Lemma 1} and {\it Lemma 3}, for $z \in S,$
%
$$ \big|f(z) - \hat f(z)\big| \leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d} \frac{\gamma k^2}{\sigma_d} \|z - x_0\|_2.$$

\noindent \textit{Proof.} Let $v = \nabla f(x_0) - c,$ where $\|v\|_2 \leq \sqrt{d} \gamma k^2 / \sigma_d$ by {\it Lemma 3}. Now
%
\begin{align*}
  \big|f(z) - \hat f(z)\big|
       &= \big|f(z) - f(x_0) - \langle c, z - x_0 \rangle \big| \\ 
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) - v, z - x_0 \rangle \big| \\
       &= \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle + \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \big| \langle v , z - x_0 \rangle \big| \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0) , z - x_0 \rangle \big| + \|v\|_2 \|z - x_0\|_2 \\
       &\leq \big|f(z) - f(x_0) - \langle \nabla f(x_0), z - x_0 \rangle \big| + \textstyle{\frac{\gamma k^2 \sqrt{d}}{\sigma_d}} \|z - x_0\|_2 \\
       &\leq \frac{\gamma \|z - x_0\|_2^2}{2} + \sqrt{d}\frac{\gamma k^2}{\sigma_d} \|z - x_0\|_2,
\end{align*}
%
where the last inequality follows from {\it Lemma 1}.
%
\qed

\section{Planning for Paper}

\begin{itemize}
\item Construct analytic function to test theoretical claims made about Delaunay.
\item Demonstrate the empirical performance of Delaunay, discuss alignment with theory.
\item Demonstrate similar empirical performance is achieved by other interpolants.
\item Argue that interpolation is not as commonly applied, but that it works well in practice.
\item Present the data for predicting real-valued outputs, an empirical study.
  \begin{itemize}
  \item Forest fire data
  \item Parkinsons
  \item https://www.kaggle.com/jsphyg/weather-dataset-rattle-package
  \end{itemize}
\item Present prediction results on each of these data sets for each algorithm.
\item Discuss the results between the regressors and interpolants.
\item Present the data for predicting function-valued outputs, an empirical study.
  \begin{itemize}
  \item VarSys large scale system configuration prediction
  \item \textit{galaxy data from Carnegie Melon?}
  \end{itemize}
\item Present prediction results on each of these data sets for each algorithm.
\item Discuss the results between the regressors and interpolants.
\item Present a summary table showing the performance differences between regerssors and interpolants.
\item Discuss the overall applicability of interpolants and encourage further application. Restate the benefits of interpolants
\item Conclude
\end{itemize}

@@

\begin{figure}
  \includegraphics[width=.5\textwidth,height=2cm]{Figures/p-hist.pdf}
  \includegraphics[width=.5\textwidth,height=2cm]{Figures/f-hist.pdf}
  \includegraphics[width=.5\textwidth,height=2cm]{Figures/h-hist.pdf}
  \caption{Histograms of Parkinsons (total UPDRS), forest fire (area), and HPC I/O (mean throughput) response values respectively. Notice that both the forest fire and HPC I/O data sets are heavily skewed.
  \vspace{-.5cm}}
  \label{fig_response_hists}
\end{figure}

@@

\begin{figure*}[htb]
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.95\textwidth,height=4cm]{Figures/eval-times.pdf}};
    \node[below=of img, node distance=0cm, yshift=1cm] {Relative Error Tolerance while Bootstrapping};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Fit Time};
  \end{tikzpicture}
  \caption{Time required to generate model fits for each technique with varying relative error tolerance during bootstrapping.
    \vspace{-.3cm}}
  \label{fig_eval_times}
\end{figure*}

@@

\begin{figure*}
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.95\textwidth,height=9.5cm]{Figures/all-performance.pdf}};
    \node[below=of img, node distance=0cm, yshift=1cm] {Relative Error Tolerance while Bootstrapping};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Signed Relative Error};
  \end{tikzpicture}
  \caption{The performance of all three techniques with varied relative error tolerance for the bootstrapping parameter. The columns are for Max Box Mesh, Iterative Box Mesh, and Voronoi Mesh, respectively. The rows are for HPC I/O, Forest Fire, and Parkinson's respectively. Notice the techniques' behavior on the Parkinson's and Forest Fire data sets, performance increases with larger error tolerance.}
  \label{fig_all_performance}
\end{figure*}

\begin{figure*}
  \begin{tikzpicture}
    \node (img)  {\includegraphics[width=0.95\textwidth,height=9.5cm]{Figures/perf-sample.pdf}};
    \node[below=of img, node distance=0cm, yshift=1cm] {Signed Relative Error in Prediction};
    \node[left=of img, node distance=0cm, rotate=90, anchor=center,yshift=-0.7cm] {Count};
  \end{tikzpicture}
  \caption{A sample of relative errors for all three techniques with optimal selections of error tolerance. The columns are for Max Box Mesh, Iterative Box Mesh, and Voronoi Mesh, respectively. The rows are for HPC I/O, Forest Fire, and Parkinson's respectively.
    \vspace{-.3cm}}
  \label{fig_perf_sample}
\end{figure*}
